% Appendix Template

\chapter{Variables and Functions in Nek5000} % Main appendix title

\label{AppendixB} % Change X to a consecutive letter; for referencing this appendix elsewhere, use \ref{AppendixX}

\lhead{Appendix B. \emph{Variables and functions in Nek}} % Change X to a consecutive letter; this is for the header on each page - perhaps a shortened title

\section{Variables}

The Nek manual provides information on many of the variables given in the .rea and SIZE
file. It is however no list of useful variables defined in other files. 
In~\tref{tab:variables} some of the variables that have been frequently used in .usr
subroutines are listed. Note that all the variables listed in this table are defined outside of SIZE, .rea and .usr 


\begin{table}
    \centering
    \begin{tabular}{l l}
        \multicolumn{2}{l}{\textbf{AVG}}\\ \hline
uavg(ax1,ay1,az1,lelt)           & Averaged values of u, similar for v,w,p\\
urms(ax1,ay1,az1,lelt)           & Variance of u, similar for v,w,p\\
vwms(ax1,ay1,az1,lelt)           & Covariance of vw\\
tavg(ax1,ay1,az1,lelt,ldimt)     & Averaged values of t and all passive scalars\\

\multicolumn{2}{l}{\textbf{GEOM}}\\ \hline
xm1(lx1,ly1,lz1,lelt)            & X-coordinates for the velocity mesh \\
xm2(lx2,ly2,lz2,lelv)            & X-coordinates for the pressure mesh\\
unx(lx1,lz1,6,lelt)              & Surface normals \\
area(lx1,lz1,6,lelt)             & Surface integral weights\\

\multicolumn{2}{l}{\textbf{INPUT}}\\ \hline

cbc(6,lelt,0:ldimt1)             & Boundary conditions of each face\\
ccurve(12,lelt)                  & Curved side character \\
curve(12,6,lelt)                 & Curved side information\\

\multicolumn{2}{l}{\textbf{PARALLEL}}\\ \hline
lglel(lelt)                      & 	Mapping from local to global element index\\
gllel(lelg)                      & 	Mapping from global to local element index\\

\multicolumn{2}{l}{\textbf{SOLN}}\\ \hline
vx(lx1,ly1,lz1,lelv)            & X-velocity\\
t(lx1,ly1,lz1,lelv,ldimt)       & Temperature and passive scalars\\
vtrans(lx1,ly1,lz1,lelt,ldimt1) & Diffusive constant to additional scalars\\
vdiff(lx1,ly1,lz1,lelt,ldimt1)  & Convective constants to additional scalars\\

\multicolumn{2}{l}{\textbf{TSTEP}}\\ \hline
istep                            & Current iteration step\\
iostep                           & Output step frequency\\
time                             & Current time\\
tstep                            & Current time step	\\
dt                               & Time step\\
dtlag(10)                        & The previous 10 time steps\\
bd(10)                           & Max 10 backward difference coeffs\\
ab(10)                           & Max 10 extrapolation coeffs (Adam-Bashforth)\\

\multicolumn{2}{l}{\textbf{WZ}}\\ \hline
zgm1(lx1,3)                      & GLL points for x,y and z directions \\

\multicolumn{2}{l}{\textbf{OTHER}}\\ \hline            
x,y,z                            & Local coordinates assigned by nekasgn()\\
ux,uy,uz                         & Local velocities assigned by nekasgn()\\
temp                             & Local temperature assigned by nekasgn()\\
nio                              & Processor node number  \\
ndim                             & Number of dimensions\\
nelv                             & Number of elements for velocity mesh\\
nelt                             & Number of elements for the t-mesh\\
pm1 (lx1,ly1,lz1,lelv)           & Pressure mapped to mesh 1\\

\end{tabular}

\caption{useful variables in Nek, the bold capital sections denote the separate files in /trunk/nek/.}

\label{tab:variables}
\end{table}

\section{Functions}

\subsection{standard calculations found in math.f or navier1.f}

\textbf{nekasgn(ix,iy,iz,ie) } Assigns the coordinates of node 
    \verb|(ix,iy,iz)| in element ie to the common variables \verb|x,y,z|.

\textbf{facind(kx1,kx2,ky1,ky2,kz1,kz2,nx1,ny1,nz1,f) } 
Assigns the index limits of a face \verb|f| with \verb|nx1,ny1,nz1| points in each spatial direction.

\textbf{zwgll(zg,wg,nx1)} Get the nx1 GLL-points and weights to \verb|zg| and \verb|wg|.

\textbf{cadd(zg,c,nx1)} Adding a constant \verb|c| to a vector \verb|zg| of length \verb|nx1|.

\textbf{cmult(zg,c,nx1)} Multiplying every element of vector \verb|zg| of length \verb|nx1| with \verb|c|.

\textbf{chsign(wrk,nx1)} change the sign of every element in vector wrk of length \verb|nx1|.

\textbf{cfill(zg,c,nx1)}  Fill vector \verb|zg| of length \verb|nx1| with the constant \verb|c|.

\textbf{rzero(zg,nx1) } Fill vector \verb|zg| of length \verb|nx1| with zeroes.

\textbf{rcopy(zg,zg2,nx1) } copy all elements from vector \verb|zg2| to vector \verb|zg|, both of length \verb|nx1|.

\subsection{Functions regarding mesh and distribution of GLL-points}

\textbf{gh\_face\_extend(x,zg,n,type,ie,v)}  The Gordon hall algorithm described in chapter~\ref{numerics}, the type variable denotes whether the 
algorithm should use vertices, edges or faces to distribute the inner GLL-points. 

\textbf{xyzlin(xl,yl,zl,nxl,nxl,nxl,ie,ifaxl)} Generate bi- or trilinear mesh.

\textbf{fix\_geom()} Routine for assuring mesh consistency and re distributing the internal  GLL-points correctly on the updated geometry.

\subsection{Additional auxiliary functions implemented for this thesis}\label{surfaceprojection}

\textbf{readsurfpro(wrksurf,iwrkelem)} Read the coordinates from \verb|surf.i| to \verb|wrksurf| and add the elements and faces in \verb|bdry.i| to \verb|iwrkelem|.

\textbf{getfaceinfo(mid,radius,wrk,ie,f)} Estimate the midpoint \verb|mid| and \verb|radius| of the face \verb|f| in element \verb|ie|. 

\textbf{createworkingarray(wrkbdry,wrksurf,mid,iter,beta,rad,f)} updating \verb|wrkbdry|, Get the points in \verb|wrksurf| within a distance \verb|rad| from the point \verb|mid|. 

\textbf{fix\_gll(ie,f)} Redistribute the GLL-points between face \verb|f| and the opposite face in element \verb|ie| to make sure that all points lie within the element.

\textbf{getface(kx1,kx2,ky1,ky2,kz1,kz2,wrk,n,ie)} assigning the values of the face in element \verb|ie| corresponding to the index limits \verb|kx1,kx2|\ldots to the array \verb|wrk(n,n,3)| .

\textbf{getsurfnorm(sn,ix,iy,iz,f,ie)} Providing the surface normal \verb|sn| at point \verb|ix,iy,iz| of element \verb|ie| and face \verb|f|. 

\textbf{calcerror(error,lambda,sn,wrk,radius)} calculate the distance \verb|error| from the initial GLL-point to a given point on the surface.  

\textbf{interp\_up(iinterp,rinterp,n,error,k)} Update interpolation points, validate if the new point \verb|k| with distance \verb|error| from the current point is closer than the existing interpolation points in \verb|iinterp|. 

\textbf{set\_new\_pt(iinterp,rinterp,ix,iy,iz,ie)} defining the position of the new GLL-point on the surface.

\textbf{getlimits(k,n,kx1,kx2,ky1,ky2,kz1,kz2)} Get the index limits \verb|kx1,kx2|\ldots corresponding to edge \verb|k| with \verb|n| GLL-points.

\textbf{setcoords(xq,yq,zq,xedge,yedge,zedge,nxl,k)} copy the updated edges \verb|xedge| to the initial edges \verb|xq|.

\textbf{getcoords(xq,yq,zq,xedge,yedge,zedge,nxl,k)} copy the node information from the initial edge \verb|xq| to \verb|xedge|.

\textbf{checkcorner(is,ix,iy,iz,kz1,kz2,ky1,ky2,kx1,kx2)} Check if the current GLL-point is a corner.

\textbf{addcorners(wrkbdry,iter,ie,f)} Add the corners of the element to \verb|wrkbdry| so that they can be used to do the projection of neighbouring points.

\subsection{Additional functions for mshconvert}
This section contains the additional functions written to supplement the script \verb|mshconvert.py|. In addition to the functions 
listed here, minor extensions were also done in the already existing functions.

\textbf{getreastart(name)} Use the first section of the file \verb|name.rea| as a template for the new \verb|.rea| file.

\textbf{getreaend(name)} Use the last section of the file \verb|name.rea| as a template for the new \verb|.rea| file.

\textbf{fixbc(name)} Fix lower case inflow boundary conditions.

\textbf{fixthermalbc(name,iftemp)} Include thermal boundary conditions if the flag \verb|temperature=true|.

\textbf{points2circ(x1,x2,x3)} Find the radius and midpoint of the circle correponding to 3 points on an arc.

\textbf{write\_surface\_file()} Write the file \verb|surf.i|.

\textbf{write\_surf\_list()} Write the file \verb|bdry.i|.

\textbf{fixSIZE()} Add or change the variables that describes the sizes of \verb|surf.i| and \verb|bdry.i| in the local \verb|SIZE| file.
